Stochastic Variance-Reduced Policy Gradient

### 2018-02-08 svrpg.tex

@@

@@

%Lemma 1
\begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
Under Assumptions \ref{ass:bounded_score} ans \ref{ass:M2}, the expected squared norm of the true gradient $\gradJ{\vtheta_t^{s+1}}$, for appropriate choices of $\alpha_t\geq0$ and $\beta_t>0$, can be bounded as follows:
\[
	\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_tV}{N\Psi_t}
	+\frac{f_tW}{B\Psi_t},
\]
	where
\begin{align*}
	&R_t^{s+1}\coloneqq \Ets[t-1]{J(\vtheta_t^{s+1}) - c_t\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}, \\
	&c_{m-1} = 0, \\
	&c_t = c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t+\frac{\alpha_t^2L^2}{B}\right)+\frac{\alpha_t^2L^3}{2B}, \\
	&\Psi_t = \alpha_t\left(\frac{1}{2}-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_tL+2\alpha_tc_{t+1}\right), \\
	&f_t = \alpha_t^2\frac{\Gamma(L+2c_{t+1})}{2},
\end{align*}
where $L=\max\left\{L_J,L_g\right\}$, \ie the greater of the Lipschitz constants from Lemmas \ref{lemma:lsmooth} and \ref{lemma:gsmooth}.

@@

\convergence*
\begin{proof}
We prove the theorem for the following values of the constants:
\begin{align*}
& \psi \coloneqq \min_t\{\Psi_t\}, 
& \zeta \coloneqq \frac{\max_t\{d_t\}V}{\psi}, 
&& \xi \coloneqq \frac{\max_t\{f_t\}W}{\psi},
\end{align*}
where $\Psi$, $d_t$ and $f_t$ are defined in Lemma~\ref{lemma:aux1}.
% The values of all other constants and the constraints on the step size are provided in Lemma \ref{lemma:aux1}.
Starting from Lemma \ref{lemma:aux1}, summing over iterations of an epoch $s$ and using telescopic sum we obtain
\begin{align*}
\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}^2}}&\leq
 \frac{\sum_{t=0}^{m-1}\left(R_{t+1}^{s+1} - R_t^{s+1}\right)}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
 & = \frac{R^{s+1}_m - R^{s+1}_0}{\psi}  + \frac{m\zeta}{N} + \frac{m\xi}{B}
\end{align*}
By using the definition of $R^s_t$ in~\eqref{E:lyapunov.function}, the fact that $c_m = 0$ and $\vtheta^{s+1}_0 = \wt{\vtheta}^s = \vtheta^s_m$, we can state that
\todopir{Why $c_m =0$? In the previous theorem  was $c_{m-1} = 0$}
\begin{align*}
        R^{s+1}_m - R^{s+1}_0 
        &= \EVV[m|s]{J(\vtheta^{s+1}_{m})- c_m \norm[]{\vtheta^{s+1}_{m} - \wt{\vtheta}^s}^2 } - \EVV[0|s]{J(\vtheta^{s+1}_{0}) - c_0 \norm[]{\vtheta^{s+1}_0 - \wt{\vtheta}^s}^2}\\
        &= \EVV[m|s]{J(\vtheta^{s+1}_{m})} - \EVV[0|s]{J(\wt{\vtheta}^s)}
        = \EVV[m|s]{J(\wt{\vtheta}^{s+1}) - J(\wt{\vtheta}^s)}
\end{align*}
Next, summing over epochs:
\begin{align}
\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}^2}}&\leq
\frac{\sum_{s=0}^{S-1}\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B} \nonumber\\
%
&\leq
\frac{\EVV[]{J(\tilde{\vtheta}^{S}) - J(\tilde{\vtheta}^{0})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}
 \label{eq:23}\\
%
&\leq
\frac{J(\vtheta^*) - J(\vtheta^0)}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}, \label{eq:24}
\end{align}
where the expectation in~\eqref{eq:23} is \wrt all the trajectories sampled in a run of Algorithm~\ref{alg:svrpg} and~\eqref{eq:24} is from the definition of $\vtheta^*$ (\ie the policy performance maximizer).
Finally, we consider the expectation \wrt all sources of randomness, including the uniform sampling of the output parameter:
\begin{align*}
\EVV[]{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} 
&=\frac{1}{T}\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} 
\leq
\frac{J(\vtheta^*) - J(\vtheta^0)}{\psi T} + \frac{\zeta}{N} + \frac{\xi}{B}.
\end{align*}

>>>>

Previous results hold for both REINFORCE and G(PO)MDP.

@@

\todopir{mention that even if for theoretical guarantees we should return a random solution, this is not a problem for us because we are evaluating the performance on-line. And is always improving}

@@

%Lemma 1
\begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
Under Assumptions \ref{ass:bounded_score} ans \ref{ass:M2}, the expected squared norm of the true gradient $\gradJ{\vtheta_t^{s+1}}$, for appropriate choices of $\alpha_t\geq0$ and $\beta_t>0$, can be bounded as follows:
\[
	\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_tV}{N\Psi_t}
	+\frac{f_tW}{B\Psi_t},
\]
	where
\begin{align*}
	&R_t^{s+1}\coloneqq \Ets[t-1]{J(\vtheta_t^{s+1}) - c_t\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}, \\
	&c_{m} = 0, \\
	&c_t = c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t+\frac{\alpha_t^2L^2}{B}\right)+\frac{\alpha_t^2L^3}{2B}, \\
	&\Psi_t = \alpha_t\left(\frac{1}{2}-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_tL+2\alpha_tc_{t+1}\right), \\
	&f_t = \alpha_t^2\frac{\Gamma(L+2c_{t+1})}{2},
\end{align*}
where $L=\max\left\{L_J,L_g\right\}$, \ie the greater of the Lipschitz constants from Lemmas \ref{lemma:lsmooth} and \ref{lemma:gsmooth}.

@@

\convergence*
\begin{proof}
We prove the theorem for the following values of the constants:
\begin{align*}
& \psi \coloneqq \min_t\{\Psi_t\}, 
& \zeta \coloneqq \frac{\max_t\{d_t\}V}{\psi}, 
&& \xi \coloneqq \frac{\max_t\{f_t\}W}{\psi},
\end{align*}
where $\Psi$, $d_t$ and $f_t$ are defined in Lemma~\ref{lemma:aux1}.
% The values of all other constants and the constraints on the step size are provided in Lemma \ref{lemma:aux1}.
Starting from Lemma \ref{lemma:aux1}, summing over iterations of an epoch $s$ and using telescopic sum we obtain
\begin{align*}
\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}^2}}&\leq
 \frac{\sum_{t=0}^{m-1}\left(R_{t+1}^{s+1} - R_t^{s+1}\right)}{\psi} + \frac{m\zeta}{N} + \frac{m\xi}{B} \nonumber\\
 & = \frac{R^{s+1}_m - R^{s+1}_0}{\psi}  + \frac{m\zeta}{N} + \frac{m\xi}{B}
\end{align*}
By using the definition of $R^s_t$ in~\eqref{E:lyapunov.function}, the fact that $c_m = 0$ and $\vtheta^{s+1}_0 = \wt{\vtheta}^s = \vtheta^s_m$, we can state that:
\begin{align*}
        R^{s+1}_m - R^{s+1}_0 
        &= \EVV[m|s]{J(\vtheta^{s+1}_{m})- c_m \norm[]{\vtheta^{s+1}_{m} - \wt{\vtheta}^s}^2 } - \EVV[0|s]{J(\vtheta^{s+1}_{0}) - c_0 \norm[]{\vtheta^{s+1}_0 - \wt{\vtheta}^s}^2}\\
        &= \EVV[m|s]{J(\vtheta^{s+1}_{m})} - \EVV[0|s]{J(\wt{\vtheta}^s)}
        = \EVV[m|s]{J(\wt{\vtheta}^{s+1}) - J(\wt{\vtheta}^s)}
\end{align*}
Next, summing over epochs:
\begin{align}
\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}^2}}&\leq
\frac{\sum_{s=0}^{S-1}\Ets[m]{J(\tilde{\vtheta}^{s+1}) - J(\tilde{\vtheta}^{s})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B} \nonumber\\
%
&\leq
\frac{\EVV[]{J(\tilde{\vtheta}^{S}) - J(\tilde{\vtheta}^{0})}}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}
 \label{eq:23}\\
%
&\leq
\frac{J(\vtheta^*) - J(\vtheta^0)}{\psi} + \frac{T\zeta}{N} + \frac{T\xi}{B}, \label{eq:24}
\end{align}
where the expectation in~\eqref{eq:23} is \wrt all the trajectories sampled in a run of Algorithm~\ref{alg:svrpg} and~\eqref{eq:24} is from the definition of $\vtheta^*$ (\ie the policy performance maximizer).
Finally, we consider the expectation \wrt all sources of randomness, including the uniform sampling of the output parameter:
\begin{align*}
\EVV[]{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} 
&=\frac{1}{T}\sum_{s=0}^{S-1}\sum_{t=0}^{m-1}\Ets{\norm[]{\gradJ{(\vtheta_t^{s+1})}}^2} 
\leq
\frac{J(\vtheta^*) - J(\vtheta^0)}{\psi T} + \frac{\zeta}{N} + \frac{\xi}{B}.
\end{align*}
